Empirical Bayes estimation of the normal inverse Gaussian prior in simultaneous-equation models

### 2019-08-13 docs/manuscript1.Rnw
updated simulations and manuscript

	\section{Introduction}
	- Motivation:
	simultaneous regressions in: network reconstruction, eQTL mapping 
	\cite[]{kpogbezan_incorporating_2019}, drug target discovery 
	\cite[]{noh_inferring_2016}.

	- Introduce data example

	- Co-data introduction

	- current literature

	- our solution: adaptive borrowing of information through empirical Bayes, 
	guided by co-data.

	\section{Model}
	\subsection{Simultaneous equations model}\label{sec:SEM}
	We have continuous measurements $y_{id}$ for observation $i=1,\dots, n$ on 
	outcome $d=1,\dots,D$. Throughout this paper we assume the $y_{id}$ to be 
	centred per outcome and let $\y_d = \begin{bmatrix} y_{1d} & \dots & y_{nd} 
	\end{bmatrix} \tr$. We predict the $y_{id}$ with features $x_{ij}$, 
	$j=1,\dots, p$, collected in $\x_i = \begin{bmatrix} x_{i1} & \dots & x_{ip} 
	\end{bmatrix} \tr$. We have characteristics of the outcomes available in the 
	form of `co-data' $\mathbf{C} = \begin{bmatrix} \mathbf{c}_1 & \dots & 
	\mathbf{c}_D \end{bmatrix} \tr$.

	We model the outcomes as a linear function of the features:
	\begin{subequations}\label{eq:linearmodel}
		\begin{align}
		y_{id} & = \beta_{0d} + \x_i \tr \bbeta_d + \epsilon_{id},\\
		\text{with } & \epsilon_{id} \sim \mathcal{N}(0, \sigma_d^2),
		\end{align}
	\end{subequations}
	where the $p$-dimensional $\bbeta_d$ are the outcome-specific feature effects.
	Note that (\ref{eq:linearmodel}) gives rise to a system of 
	$D$ linear simultaneous equations.

	\subsection{Bayesian prior model}
	We capture the uncertainty in the parameters through a Bayesian prior model. 
	We assign a flat prior to the intercept terms $\beta_{0d}$ and integrate 
	them out. The remaining parameters are endowed with the following priors:
	\begin{subequations}\label{eq:priormodel}
	  \begin{align}
      y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} 
      (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
      \beta_{jd} | \gamma_d^2, \sigma_d^2 & \sim 
      \mathcal{N} (0, \sigma_d^2 \gamma_d^2), \label{eq:betaprior} \\
      \gamma_d^2 & \sim \mathcal{IG}(\theta_d, \lambda_d), \\
      \sigma_d^2 & \sim 1/\sigma_d^3,
    \end{align}
  \end{subequations}
  where $\mathcal{IG}(\theta, \lambda)$ denotes the inverse Gaussian 
  distribution with mean $\theta$ and shape $\lambda$.
  \cite{moran_variance_2018} argue that in regression models, $\bm{\beta}_d$ 
  and $\sigma_d^2$ should be independent \textit{a priori}, i.e.
  $\beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \beta_{jd} | 
  \gamma_d^2$. However, in our case
  that is debatable. The $D$ regressions need not be on the same scale, so in
  order to jointly consider the $\bm{\beta}_d$ some 
  form of calibration is required. A straightforward way 
  of doing this is to scale the $\bm{\beta}_d$ prior with 
  $\sigma_d^2$ as in (\ref{eq:betaprior}), resulting in a 
  scale-free prior with respect to the response.

  The normal inverse Gaussian (NIG) prior model for $\beta_{jd}$ and 
  $\gamma^2_d$ in (\ref{eq:priormodel}) was introduced in 
  \cite{barndorff-nielsen_hyperbolic_1978} and is, since 
  \cite{barndorff-nielsen_normal_1997}, routinely applied in mathematical 
  finance (see, e.g., \cite{kalemanova_normal_2007}). See SM Section 
  \ref{sm-sec:nigprior} for more details on this prior.
  In Figure \ref{fig:dens_igaussian_marginalbeta}, the NIG 
  prior distribution marginalised over $\gamma_d^2$ is depicted together with 
  three other common prior choices,
  Student's $t$ (see Section \ref{sec:inversegammamodel} for more details on 
  this prior), ridge and lasso, all scaled to variance one. From this Figure 
  we see that, depending on the choice
  of hyperparameters, the NIG prior can vary from heavier tails than the lasso, 
  to more Gaussian tails like the ridge, but behaves similarly to the 
  Student's $t$ prior. Our argumentation to model the 
  $\gamma^2_d$ by an inverse Gaussian 
  distribution, as has been suggested in \cite{fabrizi_specification_2016} and 
  \cite{caron_sparse_2008}, is two-fold: (i) the NIG model is more flexible than
  the standard ridge and lasso priors (as seen from Figure 
  \ref{fig:dens_igaussian_marginalbeta}), and (ii) the NIG prior allows to model
  the mean $\theta_d$ as a 
  function of the co-data $\mathbf{c}_d$ more conveniently than the Student's 
  $t$ prior, as explained in Section 
  \ref{sec:empiricalbayes}. 
<<dens_igaussian_marginalbeta, fig.cap="Several choices of marginal $\\beta_j$ prior, scaled to $\\V(\\beta_j)=1$.", out.width="80%", fig.asp=2/3>>=
library(GeneralizedHyperbolic)
library(metRology)
# all scaled such that variance is one
sigma <- 1
theta <- 1/sigma^2
lambda <- c(0.1, 2)
lambda1 <- sqrt(2)
eta <- lambda/theta + 2
col <- c(1:6)
lty <- c(1:6)
labels <- as.expression(c(bquote("NIG, "~lambda==.(lambda[1])~", "~
                                   theta==.(theta)),
                          bquote("NIG, "~lambda==.(lambda[2])~", "~
                                   theta==.(theta)),
                          bquote("Student's t, "~lambda==.(lambda[1])~", "~
                                   eta==.(eta[1])),
                          bquote("Student's t, "~lambda==.(lambda[2])~", "~
                                   eta==.(eta[2])),
                          "ridge", "lasso"))
dprior <- function(x, lambda, theta, sigma) {
  dnig(x, 0, sigma*sqrt(lambda), sqrt(lambda/(theta*sigma)), 0)
}
dlasso <- function(x, lambda1) {
  0.5*lambda1*exp(-lambda1*abs(x))
}

ylim <- c(0, dt.scaled(0, eta[1]/2, 0, sqrt(lambda[1]*sigma^2/eta[1])))
curve(dprior(x, lambda[1], theta, sigma), -3, 3, ylab=expression(p(beta[j])), 
      xlab=expression(beta[j]), n=1000, ylim=ylim,
      col=col[1], lty=lty[1])
curve(dprior(x, lambda[2], theta, sigma), add=TRUE, n=1000, 
      col=col[2], lty=lty[2])
curve(dnorm(x, 0, 1), add=TRUE, n=1000, col=col[3], lty=lty[3])
curve(dlasso(x, lambda1), add=TRUE, n=1000, col=col[4], lty=lty[4])
curve(dt.scaled(x, eta[1]/2, 0, sqrt(lambda[1]*sigma^2/eta[1])), add=TRUE, 
      n=1000, col=col[5], lty=lty[5])
curve(dt.scaled(x, eta[2]/2, 0, sqrt(lambda[2]*sigma^2/eta[2])), add=TRUE, 
      n=1000, col=col[6], lty=lty[6])
legend("topright", legend=labels, col=col, lty=lty, 
       title="Prior", seg.len=1)
@

  A few remarks on the choice of error variance prior are justified here: 
  many authors endow error variance components with vague gamma priors. 
  \cite{gelman_prior_2006}, 
  among others, advises against this practice. The degree of `vagueness' has a 
  large influence on the posterior, while degree of `vagueness' is a difficult 
  parameter to set. This influence is especially pronounced if the likelihood is
  relatively flat, as may be reasonably expected in the large $p$, small $n$ 
  setting we are in. We therefore model the error variance with Jeffrey's 
  objective prior 
  \cite[]{jeffreys_invariant_1946}. In the derivation of our Jeffrey's prior
  for the error variance, we jointly consider an unknown mean and variance.

  - Jeffrey's prior for variance $1/\sigma^{3/2}$ (a priori independent data 
	mean and variance) or $1/\sigma^3$ (joint prior, but only mean and variance, 
	not variance and betas).

	\section{Estimation}
	\subsection{Variational Bayes}\label{sec:variationalbayes}
	The posterior corresponding to the model described in (\ref{eq:linearmodel})
	and (\ref{eq:priormodel}) is not available in closed form. We therefore 
	approximate the posterior by variational Bayes, where we force the approximate
	posterior 
	density to factorise as: $p(\bbeta_d, \sigma_d^2, \gamma_d^2) \approx 
	Q_d(\cdot) = q(\bbeta_d) \cdot q(\sigma_d^2) \cdot q(\gamma_d^2)$. For
	notational convenience, we slightly abuse notation and let $q(\cdot)$ denote
	different densities for different inputs.
	Under such a factorisation, the marginal variational posteriors that 
	minimise the Kullback-Leibler divergence of the true posterior to the 
	variational Bayes approximation \cite[]{neal_view_1998} are given by:
	\begin{align*}
    q(\bm{\beta}_d) & \overset{D}{=} \mathcal{N}_p 
    (\bm{\mu}_d, \bm{\Sigma}_d), \\
    q(\gamma^2_d) & \overset{D}{=} \mathcal{GIG}
    \(-\frac{p+1}{2}, \lambda_d \theta_d^{-2}, \delta_d\), \\
    q(\sigma^2_d) & \overset{D}{=} \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_d\).
  \end{align*}
	See SM Section \ref{sm-sec:vbderivations} for the derivations. The variational
	parameters $\bm{\mu}_d$, $\bm{\Sigma}_d$, $\delta_d$, and
	$\zeta_d$ contain cyclic dependencies and are iteratively 
	updated by:
  \begin{align*}
    \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} 
    (\X \tr \X + b_d^{(h)} \I)^{-1}, \\
    \bm{\mu}_d^{(h+1)} & = (\X \tr \X + b_d^{(h)} \I)^{-1} \X \tr \y_d, \\
    \delta_d^{(h+1)} & = a_d^{(h)} \[(\bmu^{(h+1)}_d) \tr \bmu^{(h+1)}_d +
    \trace (\bSigma_d^{(h+1)})\] + \lambda_d, \\ 
    \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + 
    (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    b_d^{(h+1)} \trace ( \bm{\Sigma}_d^{(h+1)}) + 
    b_d^{(h+1)} (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)}\bigg],
  \end{align*}
  until convergence. Here, we set
  \begin{align}
    a_d^{(h)} & =\E_{Q^{(h)}}(\sigma_d^{-2})=(n + p + 1)/(2 \zeta_d^{(h)}), 
    \nonumber\\
    b_d^{(h)} & = \E_{Q^{(h)}}(\gamma_d^{-2}) = \sqrt{\frac{\lambda_d}
    {\theta_d^2 \delta_d^{(h)}}} \frac{K_{\frac{p - 1}{2}} 
    \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)}
    {K_{\frac{p+1}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)} + 
    \frac{p+1}{\delta_d^{(h)}}, \label{eq:ratiomodifiedbessel}
  \end{align}
  where $K_{\nu}(x)$ denotes the modified Bessel function of the second kind.
  A method for fast and numerically stable calculation of ratios of such 
  functions as in (\ref{eq:ratiomodifiedbessel}) is given in SM Section 
  \ref{sm-sec:ratiosmodifiedbessels}.

  \subsection{Empirical Bayes}\label{sec:empiricalbayes}
  A full Bayesian model requires the specification of hyper parameters 
  $\bm{\theta}=\begin{bmatrix} \theta_1 & \cdots & \theta_D \end{bmatrix} \tr$ 
  and $\begin{bmatrix} \lambda_1 & \cdots & \lambda_D \end{bmatrix} \tr$. 
  These are abstract and hard to interpret parameters 
  for which we generally lack expert knowledge. They do, however, have a 
  significant influence on the shape of the posterior distribution. We 
  therefore propose to estimate these hyper parameters by empirical Bayes. 
  Simply put, empirical Bayes fits the prior model to the data and is, as such, 
  objective (up to model specification).

	The canonical method for empirical Bayes is maximisation of the marginal 
	likelihood with respect to the hyper parameters. In 
	\cite{casella_empirical_2001} the marginal likelihood is maximised by an EM 
	algorithm:
	$$
	\bm{\theta}^{(l+1)},\bm{\lambda}^{(l+1)} = \underset{\bm{\theta},\bm{\lambda}}
	{\argmax}\E_{\cdot | \Y} 
	[\log p(\Y, \B, \bgamma^2, \bsigma^2) | \bm{\theta}^{(l)},\bm{\lambda}^{(l)}],
	$$
	where $\Y = \begin{bmatrix} \y_1 & \cdots & \y_D \end{bmatrix}$,
	$\B = \begin{bmatrix} \bbeta_1 & \cdots & \bbeta_D \end{bmatrix}$, 
	$\bgamma^2 = \begin{bmatrix} \gamma^2_1 & \cdots & \gamma^2_D 
	\end{bmatrix}\tr$, $\bsigma^2 = \begin{bmatrix} \sigma^2_1 & \cdots & 
	\sigma^2_D \end{bmatrix} \tr$,
	and the expectation is with respect to the posterior. In our case, this 
	posterior is not available in closed form, which renders the expectation 
	difficult. While \cite{casella_empirical_2001} suggests to approximate the 
	expectation by a Monte Carlo sample, we propose to use the variational Bayes 
	approximation developed in Section \ref{sec:variationalbayes}:
	\begin{align*}
	  \bm{\theta}^{(l+1)},\bm{\lambda}^{(l+1)} & = 
	  \underset{\bm{\theta},\bm{\lambda}}{\argmax}\E_{Q^{(l)}} 
	  [\log p(\Y, \B, \bgamma^2, \bsigma^2)] \\
	  & = \underset{\bm{\theta},\bm{\lambda}}{\argmax} \E_{Q^{(l)}} 
	  [\log \pi (\B | \bgamma^2, \bsigma^2)] + 
	  \E_{Q^{(l)}} [\log \pi (\bgamma^2)].
	\end{align*}
	where now the expectation is with respect to the converged variational 
	posterior $Q^{(l)}=\prod_{d=1}^D Q_d^{(l)}$.

	To make use of the outcome-specific codata, we parametrise 
	$\theta_d = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ and estimate $\blambda$ and
	$\bm{\alpha}$. This allows us to include continuous co-data on the outcomes
	in an objective, data-driven manner. Additionally, it reduces the number of
	hyperparameters even further. The estimating equations are:
  \begin{align*}
    \bm{\alpha} & = \[ \mathbf{C} \tr \diag 
    (\lambda_d e_d^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \diag 
    (\lambda_d) \mathbf{1}_{D \times 1}, \\
    \lambda_d & = \[ b_d^{(l)} + e_d^{(l)} (\mathbf{c}_d \tr \bm{\alpha})^2 - 
    2\mathbf{c}_d \tr \bm{\alpha} \]^{-1},
  \end{align*}
  where $e_d^{(l)} = \E_{Q^{(l)}}(\gamma_d^2) = [b_d^{(l)} - 
  (p + 1)/\delta_d^{(l)})] \cdot \delta_d^{(l)} 
  (\theta_d^{(h)})^2/\lambda_d^{(l)}$. Solving the equations is done by 
  iteratively reweighted least squares of responses $(e_d^{(l)})^{-1}$ on 
  predictors $\mathbf{c}_d$ with weights $\lambda_d e_d^{(l)}$.

  Variaional Bayes approximations are known to underestimate posterior 
  variances \cite[]{rue_approximate_2009,consonni_mean-field_2007,
  bishop_pattern_2006,wang_inadequacy_2005}. If posterior variances are needed, 
  we suggest generating samples from the posterior after the empirical Bayes 
  iterations have converged. A Gibbs sampler to do so is described in 
  the SM Section \ref{sm-sec:gibbssampler}. Alternatively, the \texttt{R} 
  package \texttt{rstan} \cite[]{guo_rstan:_2018} allows the user to sample
  from a generic posterior. An implementation of model
  (\ref{eq:priormodel}) in \texttt{stan} code may be found on 
  \url{https://github.com/magnusmunch/cambridge}.

>>>>

	\section{Introduction}
	Recently, the dimensions of data sets have increased in three directions:
	(i) samples, (ii) variables, and (iii) outcomes. Here, we will 
	concern ourselves with directions (ii) and (iii). More spefically, we will 
	focus on the regression setting
	where we have multiple outcomes and a medium to high dimensional feature 
	space, i.e., $p \gtrsim n$. Standard methods
	underperform or even break down in such medium to high dimensional settings, 
	so new methodology is required.

	As a motivating example, we present an expression quantitative loci (eQTLs)
	detection problem (see Section \ref{sec:application}). In eQTLs detection, 
	the aim is to find a set of single nucleotide loci (SNPs) that explains a
	(part of a) gene's expression. The resulting heritability of the gene 
	expression traits is thought to be indicative of the relationship between 
	genotype and phenotype. eQTLs detection is often done for a set of genes 
	simultaneously and as such comprises a simultaneous-equations model (SEM) with
	multiple outcomes and a high dimensional feature set (the SNPs). Apart from
	eQTLs, SEM models are also applied in network reconstruction 
	\cite[]{kpogbezan_incorporating_2019, leday_gene_2017} 
	and drug target discovery \cite[]{noh_inferring_2016}, [! ref to second 
	manuscript].

	In many multi-outcome regression settings, extra information on both the 
	outcomes and 
	features is available. In our motivating eQTLs detection example, extra 
	information may come in the form of (i) gene annotation (e.g., length of the
	gene), (ii) SNP annotation (e.g., whether the SNP is in the gene or not), 
	(iii) SNP-gene annotation (e.g., distance of SNP to gene). The presented 
	method allows for the incorporation of such information to aid in the 
	estimation, by pooling information both across equations (genes) and 
	features (SNPs).

	We are not the first to propose to learn from external data 
	across equations and features. \cite{kpogbezan_incorporating_2019} uses 
	the presence or absence of edges in a normal tissue apoptosis pathway, to aid 
	in the estimation of a cancer tissue apoptosis pathway. Furhtermore, in an 
	eQTLs detection problem they include the location of the SNPs (inside
	the gene or not) to aid in the estimation. \cite{oman_minimax_2002} uses 
	empirical Bayes to learn from external data concerning both the outcomes and 
	the features in a low dimensional chemometrics dataset. Our proposed method,
	termed [! acronym here],
	extends \cite{kpogbezan_incorporating_2019,kpogbezan_empirical_2017} by 
	allowing for the inclusion of continuous external information and several 
	sources of information.

	- think of acronym

	The rest of the paper is structured as follows. In Section 
	\ref{sec:model} we introduce our model, the estimation of which is detailed
	in Section \ref{sec:estimation}. In Sections \ref{sec:simulations} and
	\ref{sec:application} we demonstrate the method in simulations and an 
  eQTLs detection problem, respectively. We end with a discussion in Section 
  \ref{sec:discussion}.

	\section{Model}\label{sec:model}
	\subsection{Simultaneous equations model}\label{sec:SEM}
	We have continuous measurements $y_{id}$ for observation $i=1,\dots, n$ on 
	outcome $d=1,\dots,D$. Throughout this paper we assume the $y_{id}$ to be 
	centred per outcome and let $\y_d = \begin{bmatrix} y_{1d} & \dots & y_{nd} 
	\end{bmatrix} \tr$. We predict the $y_{id}$ with features $x_{ij}$, 
	$j=1,\dots, p$, collected in $\x_i = \begin{bmatrix} x_{i1} & \dots & x_{ip} 
	\end{bmatrix} \tr$ and centered per feature. We have $|\bm{\alpha}|$ 
	external variables available that describes the outcome $d$, feature $j$, or
	relationship between feature $j$ and outcome $d$, collected in 
  $\mathbf{C} = \begin{bmatrix} \mathbf{c}_{1}^1 & \cdots &
	\mathbf{c}_{1}^p & \cdots & \mathbf{c}_{D}^1 & \cdots & \mathbf{c}_{D}^p
  \end{bmatrix}_{|\bm{\alpha}| \times pD} \tr$. An example of such an external
  variable from eQTL detection is the distance between SNP and gene.

	We model the outcomes $y_{id}$ as a linear function of the features $x_{ij}$:
	\begin{subequations}\label{eq:linearmodel}
		\begin{align}
		y_{id} & = \beta_{0d} + \x_i \tr \bbeta_d + \epsilon_{id},\\
		\text{with } & \epsilon_{id} \sim \mathcal{N}(0, \sigma_d^2),
		\end{align}
	\end{subequations}
	where the $p$-dimensional $\bbeta_d$ are the outcome-specific feature effects.
	Note that (\ref{eq:linearmodel}) gives rise to a system of 
	$D$ linear simultaneous equations.

	\subsection{Bayesian prior model}
	We capture the uncertainty in the parameters through a Bayesian prior model. 
	We assign a flat prior to the intercept terms $\beta_{0d}$ and integrate 
	them out. The remaining parameters are endowed with priors such that we end
	up with the following model:
	\begin{subequations}\label{eq:priormodel}
    \begin{align}
      y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} 
      \(\x_i \tr \bm{\beta}_d, \sigma_d^2\), \\
      \beta_{jd} | \gamma_{jd}^2, \sigma_d^2 & \sim \mathcal{N} 
      \(0, \sigma_d^2 \gamma_{jd}^2 \), \label{eq:betaprior} \\
      \gamma_{jd}^2 & \sim \mathcal{IG}
      \( [(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}, \lambda_{d} \), \\
      \sigma_d^2 & \sim 1/\sigma_d^3,
    \end{align}
  \end{subequations}
  where $\mathcal{IG} ([(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}, \lambda_{d})$ 
  denotes the inverse Gaussian distribution with mean 
  $[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}$ and shape $\lambda_d$.

  The normal inverse Gaussian (NIG) prior model for $\beta_{jd}$ and 
  $\gamma^2_{jd}$ in (\ref{eq:priormodel}) was introduced in 
  \cite{barndorff-nielsen_hyperbolic_1978} and is, since 
  \cite{barndorff-nielsen_normal_1997}, routinely applied in mathematical 
  finance (see, e.g., \cite{kalemanova_normal_2007}). Supplementary Material
  (SM) Section \ref{sm-sec:nigprior} contains more details on this prior.
  In Figure \ref{fig:dens_igaussian_marginalbeta}, the NIG 
  prior distribution, marginalised over $\gamma_{jd}^2$, is depicted together 
  with three other common choices,
  the Student's $t$ (see Section \ref{sec:inversegammamodel} for more details on 
  this prior), normal (ridge) and Laplace (lasso) priors, 
  all scaled to variance one. 
  From this Figure we see that, depending on the choice
  of hyperparameters, the NIG prior can vary from heavier tails than the lasso, 
  to more Gaussian tails like the ridge, but behaves similarly to the 
  Student's $t$ prior. Our argumentation to model the 
  $\gamma^2_{jd}$ by an inverse Gaussian 
  distribution, as has been suggested in \cite{fabrizi_specification_2016} and 
  \cite{caron_sparse_2008}, is two-fold: (i) the NIG model is more flexible than
  the standard ridge and lasso priors (as seen from Figure 
  \ref{fig:dens_igaussian_marginalbeta}), and (ii) the NIG prior allows to model
  the mean as a 
  function of the external variables $\mathbf{c}_{d}^j$ more conveniently than 
  the Student's $t$ prior, as explained in Section 
  \ref{sec:empiricalbayes}. 
<<dens_igaussian_marginalbeta, fig.cap="Several choices of marginal $\\beta_j$ prior, scaled to $\\V(\\beta_j)=1$.", out.width="60%", fig.asp=3/4>>=
@

  \cite{moran_variance_2018} argue that in regression models, $\bm{\beta}_d$ 
  and $\sigma_d^2$ should be independent \textit{a priori}, i.e.
  $\beta_{jd} | \gamma_{jd}^2, \sigma_d^2 \sim \beta_{jd} | 
  \gamma_{jd}^2$. However, in our case
  that is debatable. The $D$ regressions need not be on the same scale, so in
  order to jointly consider the $\bm{\beta}_d$ some 
  form of calibration is required. A straightforward way 
  of doing this is to scale the $\bm{\beta}_d$ prior with 
  $\sigma_d^2$ as in (\ref{eq:betaprior}), resulting in a 
  scale-free prior with respect to the response. Apart from the scaling,
  in this setting, the conjugate prior model is preferred
  over the non-conjugate model, because the variational Bayes approximation 
  (see Section \ref{sec:estimation}) assumes independent $\sigma_d^2$ and 
  $\bm{\beta}_d$ in the posterior, which is actually true for 
  the conjugate model. Therefore, the approximate expectation used in the 
  empirical Bayes step (see Section \ref{sec:estimation}) is more accurate.

  A few remarks on the choice of error variance prior are justified here: 
  many authors endow error variance components with vague gamma priors. 
  \cite{gelman_prior_2006}, 
  among others, advises against this practice. The degree of `vagueness' has a 
  large influence on the posterior, while degree of `vagueness' is a difficult 
  parameter to set. This influence is especially pronounced if the likelihood is
  relatively flat, as may be reasonably expected in the large $p$, small $n$ 
  setting we are in. We therefore model the error variance with Jeffrey's 
  objective prior 
  \cite[]{jeffreys_invariant_1946}. In the derivation of our Jeffrey's prior
  for the error variance, we jointly consider an unknown mean and variance. This
  joint consideration results in the somewhat unorthodox $1/\sigma^3$ Jeffrey's
  prior.

	\section{Estimation}\label{sec:estimation}
	\subsection{Variational Bayes}\label{sec:variationalbayes}
	The posterior corresponding to the model described in (\ref{eq:linearmodel})
	and (\ref{eq:priormodel}) is not available in closed form. We therefore 
	approximate the posterior by variational Bayes, where we force the approximate
	posterior 
	density to factorise as: $p(\bbeta_d, \sigma_d^2, \gamma_{1d}^2, \dots, 
	\gamma_{pd}^2) \approx 
	Q_d(\cdot) = q(\bbeta_d) \cdot q(\sigma_d^2) \cdot \prod_{j=1}^p 
	q(\gamma_{jd}^2)$. For
	notational convenience, we slightly abuse notation and let $q(\cdot)$ denote
	different densities for different inputs.
	Under such a factorisation, the marginal variational posteriors that 
	minimise the Kullback-Leibler divergence of the true posterior to the 
	variational Bayes approximation \cite[]{neal_view_1998} are given by:
	\begin{align*}
    q(\bm{\beta}_d) & \overset{D}{=} \mathcal{N}_p 
    (\bm{\mu}_d, \bm{\Sigma}_d), \\
    q(\gamma_{jd}^2) & \overset{D}{=} \mathcal{GIG}
    \(-1, \lambda_{d}[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^2, \delta_{jd}\), \\
    q(\sigma^2_d) & \overset{D}{=} \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_d\).
  \end{align*}
	See SM Section \ref{sm-sec:vbderivations} for the derivations. The variational
	parameters $\bm{\mu}_d$, $\bm{\Sigma}_d$, $\delta_d$, and
	$\zeta_d$ contain cyclic dependencies and are iteratively 
	updated by:
  \begin{align*}
    \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} 
    \[\X \tr \X + \diag(b_{jd}^{(h)})\]^{-1}, \\
    \bm{\mu}_d^{(h+1)} & = \[\X \tr \X + \diag(b_{jd}^{(h)})\]^{-1} 
    \X \tr \y_d, \\
    \delta_{jd}^{(h+1)} & = a_d^{(h)} \[(\bmu^{(h+1)}_{jd})^2 + 
    (\bSigma_d)_{jj}\] + \lambda_{d}, \\ 
    \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + 
    (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    \trace \[ \diag (b_{jd}^{(h+1)}) \bm{\Sigma}_d^{(h+1)}\] + 
    (\bm{\mu}_d^{(h+1)}) \tr \diag (b_{jd}^{(h+1)}) 
    \bm{\mu}_d^{(h+1)}\bigg],
  \end{align*}
  until convergence. Here, we set
  \begin{align}
    a_d^{(h)} & =\E_{Q^{(h)}}(\sigma_d^{-2})=(n + p + 1)/(2 \zeta_d^{(h)}), 
    \nonumber\\
    b_{jd}^{(h)} & = \E_{Q^{(h)}}(\gamma_{jd}^{-2}) = 
    \sqrt{\frac{\delta_{jd}^{(h)}}
    {\lambda_{d}[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^2}}
    \frac{K_0(\sqrt{\delta_{jd}^{(h)}\lambda_{d}}
    [(\mathbf{c}_{d}^j) \tr \bm{\alpha}])}
    {K_1(\sqrt{\delta_{jd}^{(h)}\lambda_{d}}
    [(\mathbf{c}_{d}^j) \tr \bm{\alpha}])} + 
    \frac{2}{\delta_{jd}^{(h)}}, \label{eq:ratiomodifiedbessel}
  \end{align}
  where $K_{\nu}(x)$ denotes the modified Bessel function of the second kind.
  A method for fast and numerically stable calculation of ratios of
  modified Bessel functions of the second kind, as in 
  (\ref{eq:ratiomodifiedbessel}), is given in SM Section 
  \ref{sm-sec:ratiosmodifiedbessels}.

  \subsection{Empirical Bayes}\label{sec:empiricalbayes}
  To make use of the external data, we parametrise the prior mean of the
  $\gamma_{jd}^2$ as $[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}$. 
  This allows us to
  include continuous external variables $\mathbf{c}_{d}^j$ into the model. 
  Additionally, it reduces the number of hyperparameters from $pD$ to 
  $|\bm{\alpha}| + D$. A full Bayesian model
  then requires the specification of the hyperparameters 
  $\bm{\alpha}$ and $\blambda = \begin{bmatrix} \lambda_1 & \cdots & \lambda_D 
  \end{bmatrix} \tr$. These are abstract and hard to interpret parameters 
  for which we generally lack expert knowledge. They do, however, have a 
  significant influence on the shape of the posterior distribution. We 
  therefore propose to estimate these hyperparameters by empirical Bayes. 
  Simply put, empirical Bayes fits the prior model to the data and is, as such, 
  objective Bayes (up to model specification). In our case, this results in an   
  objective and data-driven inclusion of the external data.

	The canonical method for empirical Bayes is maximisation of the marginal 
	likelihood with respect to the hyper parameters. In 
	\cite{casella_empirical_2001} the marginal likelihood is maximised by an EM 
	algorithm:
	\begin{align*}
	  \bm{\alpha}^{(l+1)},\bm{\lambda}^{(l+1)} & = \underset{\bm{\alpha},
	  \bm{\lambda}}{\argmax}\E_{\cdot | \Y} 
	  [\log p(\Y, \B, \bm{\Gamma}^2, \bsigma^2) | \bm{\alpha}^{(l)},
	  \bm{\lambda}^{(l)}] \\
	  & = \underset{\bm{\alpha},\bm{\lambda}}{\argmax} 
	  \E_{\cdot | \Y} [\log \pi (\bm{\Gamma}^2) | \bm{\alpha}^{(l)},
	  \bm{\lambda}^{(l)}],
	\end{align*}
	where $\Y = \begin{bmatrix} \y_1 & \cdots & \y_D \end{bmatrix}$,
	$\B = \begin{bmatrix} \bbeta_1 & \cdots & \bbeta_D \end{bmatrix}$, 
	$\bsigma^2 = \begin{bmatrix} \sigma^2_1 & \cdots & 
	\sigma^2_D \end{bmatrix} \tr$,
	and $\bm{\Gamma}^2 = \begin{bmatrix} \bgamma^2_1 & \cdots & \bgamma^2_D 
	\end{bmatrix}\tr$, with 
	$\bgamma^2_d = \begin{bmatrix} \gamma^2_{1d} & \cdots & \gamma^2_{pd} 
	\end{bmatrix}\tr$,
	and the expectation is with respect to the joint posterior. In our case, this 
	posterior is not available in closed form, which renders the expectation 
	difficult. While \cite{casella_empirical_2001} suggests to approximate the 
	expectation by a Monte Carlo sample, we propose to use the variational Bayes 
	approximation developed in Section \ref{sec:variationalbayes}:
	$$
	\bm{\alpha}^{(l+1)},\bm{\lambda}^{(l+1)} = 
	\underset{\bm{\alpha},\bm{\lambda}}{\argmax}\E_{Q^{(l)}} 
	[\log \pi(\bm{\Gamma}^2)],
	$$
	where now the expectation is with respect to the converged variational 
	posterior $Q^{(l)}=\prod_{d=1}^D Q_d^{(l)}$. The resulting estimating 
	equations are:
  \begin{align*}
    \bm{\alpha} & = \[ \mathbf{C} \tr \bLambda \diag 
    (e_{jd}^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \bLambda
    \mathbf{1}_{pD \times 1}, \\
    \lambda_{d} & = p \left\{ \sum_{j=1}^p b_{jd}^{(l)} + 
    \sum_{j=1}^p e_{jd}^{(l)} [(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^2 - 
    2 \sum_{j=1}^p (\mathbf{c}_{d}^j) \tr \bm{\alpha} \right\}^{-1},
  \end{align*}
  where $\bLambda=\diag(\lambda_d) \otimes \I_p$
  and $e_{jd}^{(l)} = \E_{Q^{(l)}}(\gamma_{jd}^2) = (b_{jd}^{(l)} - 
  2/\delta_{jd}^{(l)}) \cdot \delta_{jd}^{(l)}/[\lambda_{d}^{(l)}
  \left\{(\mathbf{c}_{d}^j) \tr \bm{\alpha}^{(l)}]^2\right\}$. 
  Solving the equations is done by 
  iteratively reweighted least squares.
